\documentclass[letterpaper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[shortlabels]{enumitem}
\usepackage{tocloft}
\usepackage{hyperref}

\hypersetup{
	colorlinks=true,
	linkcolor=black,
	filecolor=magenta,      
	urlcolor=black,
	pdfauthor="Lily Jade Joyce"
	hyperindex=true
	bookmarks=true
	bookmarksopen=true
}

\renewcommand{\thesection}{\Roman{section}}
\addtolength{\cftsecnumwidth}{10pt}

\title{Rensselaer Polytechnic Institute Graduate Council Bylaws}
\date{ }
\begin{document}
\begin{titlepage}
	\maketitle
	\thispagestyle{empty}
\end{titlepage}

\newpage %somewhat hacky way to make the page numbers start on the right page
\tableofcontents
\thispagestyle{empty}
\newpage
\clearpage
\pagenumbering{arabic}

\section{Preamble}
\noindent These ByLaws supersede all previous ByLaws of the Graduate Student Council, effective upon adoption of the Graduate Student Council.

\section{Purpose}
\begin{enumerate}[A.]
	\item The Graduate Student Council (hereafter, the Council) is the elected representative body of the Graduate Students and the graduate legislative branch of the Rensselaer Union.
	\item The Council shall be responsible for initiating all legislative action within the Rensselaer Union pertaining only to Graduate Students and not specifically within the domain of the Student Senate or the Executive Board.
	\item The Council shall act as an advisory board to the Graduate School.
	\item The Council shall ensure adequate representation to other governing bodies by appointing members to other committees as prescribed by the Rensselaer Union Constitution.
\end{enumerate}

\section{Eligibility}
\begin{enumerate}[A.]
	\item Students who are enrolled in the Graduate School and who are pursuing a graduate degree at Rensselaer shall be eligible to run for and hold office on the Graduate Student Council. Students who do not meet these requirements at any time shall not be considered to be on the Council and their office shall be considered vacant.
	\item Appointments made by the Graduate Student Council shall be subjected to the same eligibility requirements
\end{enumerate}

\section{Membership}
\begin{enumerate}[A.]
	\item Only students meeting the above eligibility requirements in Article III may run for and hold office on the Graduate Student Council.
	\item  The voting membership of the Graduate Student Council shall consist of all Graduate Senators and all Graduate Representatives.
	\begin{enumerate}[1.]
		\item Seven members shall be elected as Graduate Senators to serve on the Student Senate in addition to the Council.
		\item Six members shall be elected as Graduate Representatives at­large.
		\item One additional Graduate Representative shall be elected by the Black Graduate Student Association.
		\item The Council shall make it possible for each school of Rensselaer to be represented on the Council by at least one Graduate Senator or Graduate Representative.
		\begin{enumerate}[1.]
			\item For the purposes of this document, “school of Rensselaer” shall be one of:
			\begin{enumerate}[1.]
				\item Information Technology
				\item Lally School of Management and Technology
				\item School of Architecture
				\item School of Engineering
				\item School of Humanities and Social Sciences
				\item School of Science
			\end{enumerate}
		\end{enumerate}
	\end{enumerate}
	\item If the At­-large Council membership is full (all Graduate Senators and all at­-large Graduate Representatives have been elected), and a school of Rensselaer is not represented on the council, then, each year, an additional at­-will Graduate Representative seat will be opened for interested applicants from the unrepresented school. These at­-will open seats will not be open by default: the requirement for the open seats will be determined not before the completion of Grand Marshal Week elections each year, and the open seat shall close at the conclusion of each term. Once opened, the seat will be filled according to the standard procedure for vacancies as detailed in Article V, Section E.
	\item The non­voting membership of the Graduate Student Council shall consist of:
	\begin{enumerate}[1.]
		\item The Dean of the Graduate School or their designee.
		\item All representatives to other bodies appointed by the Council, if not already voting members.
		\begin{enumerate}[1.]
			\item A student may act as a representative to more than one body.
		\end{enumerate}
		\item The Grand Marshal and the President of the Union, if they meet the eligibility criteria stated in Article III above.
		\item The president of any other graduate organization at Rensselaer that: advocates for graduate interests, or holds or plans social events. These graduate organizations include, but are not limited to: the School of Science Council, the Center for Biotechnology and Interdisciplinary Studies Council, and the MANE Student Advisory Council.
	\end{enumerate}
\end{enumerate}

\section{Electrons and Terms of Office}
\begin{enumerate}[A.]
	\item Council elections shall be held during Grand Marshal Week and shall be run by the Student Senate.
	\begin{enumerate}[1.]
		\item Students shall take office immediately upon announcement of election results, at which time the term of office for previous members shall expire.
	\end{enumerate}
	\item A Council member’s term of office shall end if, prior to Grand Marshal Week, the Council member:
	\begin{enumerate}[1.]
		\item resigns
		\item is removed
		\item graduates or otherwise leaves Rensselaer
	\end{enumerate} 
	\item Removal of a member from the Council, or from any appointed or elective position, shall require concurrence of two-­thirds of the remaining voting membership of the Council.
	\item Resignation or removal from the Student Senate shall constitute removal from the Council. A senator removed in this way may apply for a position as Member­-at-­Large, if such position is vacant.
	\item Vacancies on the Council and appointments to other bodies may be filled by appointment with the approval of two­-thirds of the voting membership of the Council.
	\begin{enumerate}[1.]
		\item Applications must be solicited publicly at least ten days in advance of any appointment.
		\item The Council shall offer to interview all applicants for a position prior to filling that position.
		\item An appointment shall be valid for the remainder of the term for office.
	\end{enumerate}
	\item In the event of the death of a Council member before the natural end of their term of office, said office shall be held open out of respect for one month.
\end{enumerate}


\section{Officers}
\begin{enumerate} [A.]
	\item As soon as practical, following the annual elections, the newly-­elected Graduate Student Council shall meet to elect from its voting membership a President, a Vice President, a Treasurer and a Secretary, whose duties shall include, but are not limited to, those outlined below.
	\begin{enumerate}[1.]
		\item  The previous Council President shall call this meeting, and preside over the election of the new president. The new president shall then preside over the election of all other officers.
	\end{enumerate}
	\item The President shall
	\begin{enumerate}[1.]
		\item be one of the officers of the Rensselaer Union.
		\item act as primary spokesman for the Council.
		\item preside over all Council meetings, or appoint a President Pro Tempore to do so if unable to attend.
		\item act as representative to the Board of Trustees, or, if unable to do so, appoint another member of the Council to this position.
		\item act as representative to the Rensselaer Alumni Association Board of Trustees.
	\end{enumerate}
	\item The Vice President shall
	\begin{enumerate}[1.]
		\item act as President in their absence, if no President Pro Tempore has been designated.
		\item coordinate the operations of the Council not provided elsewhere in these By­Laws.
	\item monitor the activities of the various committees.
	\item be responsible for maintaining electronic communication channels of the Council.
	\end{enumerate}
	\item The Treasurer shall
	\begin{enumerate}[1.]
		\item handle all monies derived from the Graduate Fee, and all other income.
		\item handle all purchases and other monetary transactions of the Council.
		\item authorize members to charge against the Council's account for amounts less than \$50. Expenditures exceeding this amount require a majority approval of the Council.
		\item keep a record of all Council expenses.
		\item make monthly reports to the Council of its financial status.
	\end{enumerate}
	\item The Secretary shall
	\begin{enumerate}[1.]
		\item take minutes to all meetings and distribute them to Council members.
		\begin{enumerate}[1.]
			\item Minutes shall be reviewed and approved by a majority vote of the Council.
		\end{enumerate}
		\item be responsible for advertising meetings and events of the Council.
	\end{enumerate}
\end{enumerate}

\section{Meetings}
\begin{enumerate}[A.]
	\item Meetings shall be held at least once per month during normal
	academic periods.
	\begin{enumerate}[1.]
		\item Telephone and electronic voting is permitted as needed between
		meetings. Any decisions made by a phone or vote must be reviewed
		at the next meeting. Email or other electronic messaging services
		are both permissible options for electronic voting.
	\end{enumerate} 
	\item The President shall call meetings and publicly post the agenda no less
	than one business day prior to the meeting.
	\item Meetings are open to all students unless closed by a majority vote of
	the Council.
	\item Council members shall be notified at least three days in advance of a
	regular meeting.
	\item Quorum is necessary to conduct business. Quorum shall consist of a
	majority of the active voting membership of the Council. The
	President, or their designee, must be present for quorum to be in
	effect.
	\item All matters, unless specified otherwise by these by-laws, shall be
	decided by a majority of those voting members present.
	\item An emergency meeting may be called at the request of any Council
	member with the consent of the majority of voting members.
	\begin{enumerate}[1.]
		\item Emergency meetings must be held within three days.
		\item It shall be the responsibility of the requesting member to obtain
		majority consent and to give all voting members at least 24 hours
		notice of the meeting.
		\item The requesting member, or their designee, shall preside over
		the meeting.
	\end{enumerate}
\end{enumerate}

\section{Graduate Student Council \\ Responsibilities}
\begin{enumerate}[A.]
	\item The Graduate Student Council shall have duties and responsibilities
	including, but not limited to, those listed below.
	\begin{enumerate}[1.]
		\item assist and advise the Student Senate on academic issues.
		\item work with the Institute to address issues related to graduate
		academics and Teaching and Research Assistants.
		\item coordinate admissions activities with the Graduate School and
		Graduate Admissions office.
		\item act as liaison and coordinate communications between the Council
		and prospective Graduate Students.
		\item include and advise the liaisons to the
		\begin{enumerate}[1.]
			\item Center for Career and Professional Development
			\item Event Planning Committee
			\item Graduate Research Symposium Taskforce
		\end{enumerate}
		\item coordinate social functions for the Graduate Student Body.
		\item nominate a graduating graduate student to speak at
		commencement.
	\end{enumerate}
	\item The Council shall have the authority to create ad-hoc committees, as deemed necessary, by a majority vote.
	\begin{enumerate}[1.]
		\item Any Graduate Student may serve on a Council committee.
		\item Committee chairs shall be appointed by the Council President.
		\item Each committee shall be responsible for maintaining records of its
		activities.
		\begin{enumerate}[1.]
			\item Committee chairs shall make a report of committee activities to the Council upon request.
		\end{enumerate}
	\end{enumerate} 
\end{enumerate}

\section{Amendments}
\begin{enumerate}[A.]
	\item Amendments to the Graduate Student Council By-Laws must be proposed at least one week before their adoption by the Graduate Student Council.
	\begin{enumerate}[1.]
		\item Adoption of an amendment by the Council shall require a two-thirds
		majority of the voting membership of the Council.
		\item After adoption, the Graduate Council shall present any changes to the Student Senate in a timely manner.
	\end{enumerate}
\end{enumerate}
\vspace{2cm}
\begin{center}
	For questions about the By-Laws of the Graduate Student Council, \\
	please e-mail us at \href{mailto:grad-council@rpi.edu}{grad-council@rpi.edu}.
\end{center}
\end{document}
